---
title: "Using Choropleth Maps for Country-level Statistics"
author: "Kelly Ng"
date: '2022-04-05'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

I need the following packages:
```{r}
library(countrycode)
library(readxl)
library(sf)
library(tidyverse)
library(tmap)
```

1. Importing data: 
```{r}
gdp_per_cap <- read_excel(
  "API_NY.GDP.PCAP.KD_DS2_en_excel_v2_3731742.xls",
  skip = 3
)
life_exp <- read_excel(
  "API_SP.DYN.LE00.IN_DS2_en_excel_v2_3731513.xls",
  skip = 3
)

# Next function can pivot any of the World Bank tibbles
pivot_longer_wb <- function(tb, name) {
  tb |>
    select(-starts_with("Indicator")) |>
    pivot_longer(
      matches("^\\d{4}$"), # String with exactly 4 digits
      names_to = "year",
      values_to = {{ name }},
      names_transform = list(year = as.integer)
    )
}
wb <-
  Hmisc::llist(life_exp, gdp_per_cap) |>
  imap(pivot_longer_wb) |>
  reduce(left_join, by = c("Country Name", "Country Code", "year")) |>
  semi_join(codelist, by = c("Country Code" = "iso3c"))
```

(2) Import the country borders as `sf` objects. Because the boundaries are 
represented by polygons with many vertices, plotting the data with tmap later 
on would be slow. Instead of importing directly with `read_sf()`, use the 
function, called `import_and_simplify()`, instead. It uses the function 
`ms_simplify()` from the **rmapshaper** package to reduce the number of 
vertices.
```{r}
import_and_simplify <- function(zip_file) {
zip_file |>
unzip() |>
str_subset(".shp$") |>
read_sf() |>
rmapshaper::ms_simplify()
}

country_borders <- import_and_simplify("wb_countries_admin0_10m.zip")
```

(3) Norway and France have invalid country codes in the column `ISO_A3`. 

Therefore, I changed them to `NOR` and `FRA` respectively.
```{r}
country_borders <- 
country_borders |>
  mutate(ISO_A3 = case_when(
    FORMAL_EN == "Kingdom of Norway" ~ "NOR",
    FORMAL_EN == "French Republic" ~ "FRA", 
    FORMAL_EN == "French Polynesia" ~ "FRA", 
    FORMAL_EN == "Territory of the French Southern and Antartctic" ~ "FRA", 
    TRUE ~ ISO_A3
  ))
```


(4) With the help of the files listed above (in the section called ‘Data’), 
lets create an sf object called countries that contains the columns:
- name
- code (with ISO 3166-1 alpha-3 country code)
- gdp_per_cap (only for 2015)
- life_exp (only for 2015)
- geometry (use the sf object you created in the previous task)

I will include as many countries as possible. However, exclude any countries for 
which the geometry is empty (i.e. unknown) because empty geometries would 
trigger a warning by the tmap package later on. You can find empty geometries in
the geometry column by using st_is_empty(countries)).
```{r}
countries <-
  country_borders |>
  inner_join(
    wb,
    by = c("ISO_A3" = "Country Code")
  ) |>
  filter(year == 2015) |>
  select(
    name = "Country Name",
    code = "ISO_A3",
    gdp_per_cap, 
    life_exp, 
    geometry
  ) |>
  print()

any(st_is_empty(countries))
```

(5) Lets look at a quick-and-dirty visualisation made with **tmap** that shows 
the geometries stored in the `countries` object. 
```{r}
tm_shape(countries, projection = "+proj=eck4") + 
  tm_polygons()
```

Are any countries or land masses missing?

Antarctica, Western Sahara, Ladakh, and Arunachal Pradesh is missing. 

6) I will add the missing polygons to the map by:

  (a) importing and simplifying the World Bank geospatial data for land masses 
with `import_and_simplify("wb_land_10m.zip")`.

  (b) plotting the land masses in grey as base layer and the country borders in 
countries as top layer.
```{r}
land_masses <- import_and_simplify("wb_land_10m.zip")

tm_shape(land_masses, projection = "+proj=eck4") +
  tm_polygons(col = "red") +
  tm_shape(countries, projection = "+proj=eck4") +
  tm_polygons()
```

(7) Because I want to make choropleth maps, I use an equal-area map projection. 
Therefore, I will find the current projection of the map. 
```{r}
st_crs(land_masses)$proj4string
```

(8) Mapping the wealth and health of countries in 2015 using choropleth maps:
```{r}
tm_shape(land_masses, projection = "+proj=eck4") +
  tm_polygons() +
  tm_shape(countries, projection = "+proj=eck4") +
  tm_polygons(
    col = c("gdp_per_cap", "life_exp"),
    title = c(
      "GDP per capita\n(US$, PPP 2015)",
      "Life expectancy at birth\n(years)"
    ),
    palette = list("OrRd", "BuGn"),
    breaks = list(
      c(
        -Inf,
        1000,
        2000,
        5000,
        10000,
        20000,
        50000,
        Inf
      ),
      c(
        -Inf,
        60,
        65,
        70,
        75,
        80,
        85,
        Inf
      )
    ),
    labels = list(
      c(
        "Less than 1,000",
        "1,000 to 2,000",
        "2,000 to 5,000",
        "5,000 to 10,000",
        "10,000 to 20,000",
        "20,000 to 50,000",
        "50,000 or more"
      ),
      c(
        "Less than 60",
        "60 to 65",
        "65 to 70",
        "70 to 75",
        "75 to 80",
        "80 to 85",
        "85 or more"
      )
    )
  ) +
  tm_layout(
    main.title = "Wealth and Health by Country in 2015", # https://stackoverflow.com/questions/61355422/how-to-put-title-of-a-map-outside-of-the-panel-tmap-package
    inner.margins = c(0.01, 0.3, 0.01, 0.01),
    bg.color = "lightblue",
    earth.boundary = TRUE,
    space.color = "white",
    legend.title.size = 1, # https://stackoverflow.com/questions/32890762/how-to-manipulate-tmap-legend
    legend.text.size = 0.8,
    main.title.size = 1.5,
    legend.position = c("left", "top"),
    frame = FALSE # https://stackoverflow.com/questions/58122122/remove-customize-facets-borders-in-tmap-tm-facets
  ) +
  tm_text(
    "code",
    size = "AREA",
    root = 5,
    remove.overlap = TRUE,
    size.lim = c(0.3, 0.5)
  ) +
  tm_credits(
    text = c("", "Source: World Bank"),
    size = 0.8,
    position = c(0.01, 0.02)
  )
```

(9) What can one conclude from the maps?

A reader can conclude that the wealth and health of countries in Africa are low compared to those of the rest of the world. Similarly, their life expectancy is also generally lower compared to those in other continents. Given this, readers can conclude that wealth and health and life expectancy have a positive relationship with one another, especially given that countries in North America have both high wealth and health and life expectancy. Furthmore, Latin American countries are around the middle range in terms of wealth and health. For Asian countries, they are generally heterogenous; Afghanistan is at the low end of the spectrum, whereeas Japan and South Korea are both wealth and live longer lives. Overall, we can observe that there is a correlation between GDP per-capita and average life expectancy. 

(10) What are the advantages and disadvantages of this visualisation in comparison to figure 1?

The advantage of this visualization compared to figure 1 is that the points representing each country do not overlap with one another, making it easier to see the data for a single country. However, it becomes harder to compare each country with one another as the shades representing different categories are quite similar. Furthermore, while figure 1 plots these countries on the x and y axes where every number is represented, the countries in the choropleth map are put into categories. Hence, one cannot tell the exact value for GDP and life expectancy that a country has. Another disadvantage is that since we have two different maps, readers have to shift their attention between the two to observe a ccorrelation between GDP per-capita and life expectancy. A way to solve this problem would be if we had an interactive map rather than a printed one, where viewers can toggle between two variables.